Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-01-27 paper/main.tex
.

\subsection{Choosing a train-test splitting scheme}
In general, drug-target interaction prediction is the task of accurately predicting, whether for a given drug and a given protein there is a biological interaction within the target organism. Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model. However, when building the train-test split over compound-protein pairs for building the actual model, there are the following three options:

\begin{enumerate}
	\item Build split over drugs
	\item Build split over drug-target pairs
	\item Build split over proteins
\end{enumerate} 
In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION). As there are hopefully many more drugs to discover, the drug split scheme both emphasizes the drug repurposing idea, by applying unseen compounds to existing targets, but also benefits from more complicated drug representations, leading to tremendous results. This performance gain is based on minor variations among large groups of pharmaceuticals, that are easy to acquire. The second scheme has knowledge on all drugs and all proteins, and is thus prone to overfitting and the same development bias. Eventually, as there only limited drug-targets \citep{Overington2006}, predicting per protein is rather counter-intuitive. As it is hard to generalize over proteins representations, we aim at reaching similar performances for both drug and protein splitting schemes. \\

In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION). The first is more relevant for novel drugs, as it is much more likely to test a new compound than a innovative protein. However, it lies in the very nature of the used datasets, making the prediction for new drugs much easier. Thus, drugs are often built by minor variations of existing drugs, thus leading to no deviations in the functional group of that very compound (CITATION/EXAMPLE). When distributed over both train and test split, the models do not perform inductive inference and generalize, but rather implement transductive inference by just predicting the recently seen structures. Hence, when entirely new molecules are seen, the models perform much worse. \\
The same applies to splits of drug-target pairs, as all drugs were already seen, and novelty cannot be coped with.\\
As mentioned in the introduction it is quite difficult to learn suitable features from proteins. In general, attempts search for motifs in the protein sequences under usage of convolutional neural networks and filters, which is more suitable for tasks like protein function prediction, than for for drug-target interaction prediction, and lack a more in-depth hypothesis on the protein side, while investing in refined drug features. \\
Thus, building splitting over proteins is the most challenging of the three options. \\

>>>>

\subsection{Choosing a train-test splitting scheme}

\todo[inline]{List a number of general challenges first, then refer
  back to them: very similar drugs are designed to target the same
  proteins, minor variations, leading to a kind of ``hidden''
  duplication of drugs in the training/testing set; some proteins act
  like hubs and interact with many drugs (making them easy and
  reliable predictors); shared indications of drugs are usually
  because they target the same proteins by design leading to some
  trivial predictions; some proteins are more ``valuable'' targets due
  to their disease involvement (such as cancer, meaning more drugs are
  developed for targeting those); }

In general, drug-target interaction prediction is the task of accurately predicting, whether for a given drug and a given protein there is a biological interaction within the target organism. Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model. However, when building the train-test split over compound-protein pairs for building the actual model, there are the following three options:

\begin{enumerate}
	\item Build split over drugs
	\item Build split over drug-target pairs
	\item Build split over proteins
\end{enumerate} 
In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION). As there are hopefully many more drugs to discover, the drug split scheme both emphasizes the drug repurposing idea, by applying unseen compounds to existing targets, but also benefits from more complicated drug representations, leading to tremendous\todo{???} results. This performance gain is based on minor variations among large groups of pharmaceuticals, that are easy to acquire\todo{evidence}. The second scheme has knowledge on all drugs and all proteins, and is thus prone to overfitting and the same development bias. Eventually, as there only limited drug-targets \citep{Overington2006}, predicting per protein is rather counter-intuitive. As it is hard to generalize over proteins representations, we aim at reaching similar performances for both drug and protein splitting schemes. \\

In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION). The first is more relevant for novel drugs, as it is much more likely to test a new compound than a innovative protein. However, it lies in the very nature of the used datasets, making the prediction for new drugs much easier. Thus, drugs are often built by minor variations of existing drugs, thus leading to no deviations in the functional group of that very compound (CITATION/EXAMPLE). When distributed over both train and test split, the models do not perform inductive inference and generalize, but rather implement transductive inference by just predicting the recently seen structures. Hence, when entirely new molecules are seen, the models perform much worse

The same applies to splits of drug-target pairs, as all drugs were already seen, and novelty cannot be coped with.

As mentioned in the introduction it is quite difficult to learn suitable features from proteins. In general, attempts search for motifs in the protein sequences under usage of convolutional neural networks and filters, which is more suitable for tasks like protein function prediction, than for for drug-target interaction prediction, and lack a more in-depth hypothesis on the protein side, while investing in refined drug features. \\
Thus, building splitting over proteins is the most challenging of the three options. \\
